Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2017-09-09 stp/paper_aaai/main.tex
fixed the typo with the serene name and a just couple more typos in the appendices so the tables don't look hyperlong, and typos in the mzn model

\begin{table}[h]
	\centering
	\caption{Feature extractors that produce a scalar value}
	\begin{tabular}{p{3cm}p{4.5cm}}
	Feature name & Description \\ \hline
	num-unique-vals & A simple count of unique entries of a column\\
	\rowcolor{Gray}
	prop-unique-vals & The proportion of unique entries of a column. \\
	\rowcolor{Gray}
	prop-missing-vals & The proportion of all missing/empty entries from all entries in a column.\\
	ratio-alpha-chars & The average of the proportion of alphabetic characters from each entry.\\
	\rowcolor{Gray}
	prop-numerical-chars & The average of the proportion of numeric characters from each entry.\\
	prop-whitespace-chars    & The average of the proportion of whitespace characters from each entry.\\
	\rowcolor{Gray}
	prop-entries-with-at-sign& The proportion of entries with an '@' sign. \\
	prop-entries-with-hyphen & The proportion of entries with a '-'. \\
	\rowcolor{Gray}
	prop-range-format numerical range format  & The proportion of entries which follow some ([0-9]+)-([0-9]+)\\

	is-discrete              & A binary indicator if the column entries are discrete. \\
	\rowcolor{Gray}
	entropy-for-discrete-values & The entropy of the entries of a column.\\
	shannon-entropy          & Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'\\
	\end{tabular}
\end{table}

@@

	\begin{tabular}{p{2.5cm}p{5cm}}
	Feature name & Description \\ \hline
inferred-data-type      & Produces a vector which contains the inferred data type.  
Only a single element is non-zero in the vector (i.e. 1-of-K format).  
The set of inferred types are in float, integer, long, boolean, date, time, datetime, string. \\
	\rowcolor{Gray}
char-dist-features      & Produces a vector for character distribution of column values;
considered characters are printable characters
%                        0123456789abcdefghijklmnopqrstuvwxyz
%                        ABCDEFGHIJKLMNOPQRSTUVWXYZ!"
%                        #$%&\'()*+,-./:;<=>?@[\\]^_`{|}~ \t\n\r\x0b\x0c
\\
stats-of-text-length    & Computes the mean, median, mode, min, max of string lengths from a columns entries.\\
	\rowcolor{Gray}
stats-of-numerical-type & Computes the mean, median, mode, min, max of entries from a column of numerical type.\\

prop-instances-per-class-in-knearestneighbours         & This finds the k columns from the training set with most similar names.
It then produces a vector of class proportions from these k neighbours.
Parameters:
"num-neighbours": The number of neighbours to find (k).\\
	\rowcolor{Gray}
mean-character-cosine-similarity-from-class-examples   & For some test instance t, this computes the character distribution 
from this instance's entries.  It then produces a vector of the average
cosine similarity of t with training instances of each class.\\
min-editdistance-from-class-examples                   & For some test instance t, this produces a vector of the minimum column name
edit distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing edit distances of words; 
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

	\rowcolor{Gray}
min-wordnet-jcn-distance-from-class-examples            & For some test instance t, this produces a vector of the minimum column name
JCN WS4J distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing wordnet distances of words;
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

min-wordnet-lin-distance-from-class-examples            & For some test instance t, this produces a vector of the minimum column name
LIN WS4J distance between t and all training instances of each class.
Parameters:
"cache-size": Size of cache for storing wordnet distances of words; 
"max-comparisons-per-class": Max number of training instances to compare
against for each class.\\

@@

\clearpage\newpage
\subsection{\minizinc{} Model \label{Ann:MZ}}
\renewcommand{\lstlistingname}{Model}
\begin{minipage}{\textwidth}
%\begin{figure}[!h]
	\begin{lstlisting}[caption={\minizinc{} model (part 1 of 2)},captionpos=b]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% MiniZinc model to solve the Steiner Tree Problem arising %%
%% in the Rel2Onto problem.                                 %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%% Steiner Tree Problem
int: nbV;                               % Nb of nodes in alignment graph
int: nbE;                               % Nb of edges in alignment graph
set of int: nodes = 1..nbV;    
set of int: edges = 1..nbE;
set of int: cnodes;                     % Class nodes in alignment graph
set of int: dnodes;                     % Data nodes in alignment graph
set of int: anodes;                     % Attribute nodes in alignment graph (normally none)
array [nodes] of var bool: vs;          % Is a node of the alignment graph in the solution?
array [edges] of var bool: es;          % Is an edge of the alignment graph in the solution?
array [nodes,edges] of bool: adjacent;  % adjacent[n,e] <=> node n is an endnode of edge e
array [edges] of nodes: heads;          % endpoint 1 of each edge (the graph is undirected)
array [edges] of nodes: tails;          % endpoint 2 of each edge
array[1..2,edges] of int: endpairs = array2d(1..2,edges,tails++heads);
array [edges] of int: ws;               % Weights of edges in alignment graph
var 0..sum(ws): w;                      % Weight of the solution corresponding to the alignment part
constraint steiner_tree(vs,es,adjacent,endpairs,w,ws);  % Steiner Tree constraint
constraint forall (d in dnodes) 
                   (vs[d] <-> sum(e in edges where adjacent[d,e])(es[e]) == 1);% Only 1 edges per data node

%%%%%%% Matching Problem
int: nbA;                                           % Nb of attribute nodes
set of int: atts = 1..nbA;                 
array [atts] of set of dnodes: attribute_domains;   % Data nodes to which each attribute can attach (usually all) 
constraint forall(att in anodes) (vs[att] == true); % Attribute nodes (normally none) are in the graph.
array [atts] of var dnodes: match;                  % Match between Attributes and Data nodes
array[atts,dnodes] of int: match_costs;             % Weight of match edges
constraint forall (a in atts) (match[a] in (attribute_domains[a]));   % Must match someone in the domain
constraint forall (a in atts) (vs[match[a]]);       % The data node matched must be in the subgraph of the alignment
constraint alldifferent(match);                     % Exact match
var 0..sum(a in atts)(sum(n in dnodes)(match_costs[a,n])): wm;  % Total cost from matching part :
constraint wm = sum(a in atts) (match_costs[a,match[a]]);

	\end{lstlisting}
%	\caption{\minizinc{} model (1/2).}
%\end{figure}
\end{minipage}
\clearpage
\newpage
\begin{minipage}{\textwidth}
%\begin{figure*}[t]
	\begin{lstlisting}[caption={\minizinc{} model (part 2 of 2)},captionpos=b]
%%%%%%% Unknown nodes
int: nb_unknown_nodes;                 % Number of unknown data nodes
set of int: unknowns = 1..nb_unknown_nodes;
array [int] of int: all_unk_nodes;     % List of all unknown nodes (first two elements are class nodes, by convention)
int: node_root = all_unk_nodes[1];     % Root node if unknown class node is used.
int: node_unk = all_unk_nodes[2];      % Unknown class node
constraint if nb_unknown_nodes > 0 then
    vs[node_unk] <-> vs[node_root]     % If class node "Unknown" is used, class node "Root" is used as well
 /\ vs[node_root] <-> sum(e in edges where adjacent[node_root,e])(es[e]) in {1,2}
else true %Ignore constraint
endif;

%%%%%%% Patterns
int: nb_patterns;                     % Number of patterns
set of int: PATTS = 1..nb_patterns;
array[PATTS] of int: patterns_w;      % Support of patterns
array[PATTS] of set of int: patterns; % The actual patterns
array[PATTS] of var bool: c_p;        % Is pattern chosen
constraint forall (p in PATTS) ((sum (e in patterns[p])   % All edges of a pattern are used <-> pattern is used
							                  	(es[e]) == card(patterns[p])) <-> c_p[p]);
var 0..sum(patterns_w): wp;           % Total prize collected from patterns : 
constraint if nb_patterns = 0 then wp = 0 else wp = sum(p in PATTS) (c_p[p] * patterns_w[p]) endif;

%%%%%%% Final objective function 
var int: obj = w + wm - wp;           % Total cost of semantic model - prize collected from patterns

@@

ann: cheap_edges_fill_pattern;        % Custom search strategy for patterns
ann: custom(ann, array[int] of int, array[int] of var bool);
array[int] of ann: cefp = 
if nb_patterns == 0 then [] 
else [custom(cheap_edges_fill_pattern,[nbE]++[nb_patterns]++ws++[ card(patterns[p]) | p in PATTS]++patterns_w++array1d([ e | p in PATTS, e in patterns[p]]),es)] endif;

% The solver will use these strategies in order:
% 1) First try using the cheapest matches, until a match is found
% 2) If there are patterns, use a specific search called "cheap_edges_fill_pattern". (Implemented in Chuffed only)
%    It will start adding edges in increasing order of cost, while trying to fill patterns that are "worth" filling.
%    E.g. We already set es[x] = true for x in {e1,e2,e3}; if adding ws[e5] = 4 and ws[e7] = 6, but
%    {e1,e2,e3,e7} form a pattern p of patterns_w[p] = 3, then ws[e7] - patterns_w[p] <  ws[e5], so e7 is selected first
%    If there are no patterns, this strategy is ignored.
% 3) Try to eliminate edges by decreasing cost: it will eliminate the more expensive edges first, until
%    the strictly necesary cheap edges are added by the propagators. This is used only when there are no 
%    patterns, otherwise (2) already selects the edges.

>>>>

\begin{table}[h]
	\centering
	\caption{Feature extractors that produce a scalar value}
	\begin{tabular}{p{3cm}p{4.5cm}}
	Feature name & Description \\ \hline
	num-unique-vals & A simple count of unique entries of a column\\
	\rowcolor{Gray}
	prop-unique-vals & The proportion of unique entries of a column. \\
	\rowcolor{Gray}
	prop-missing-vals & The proportion of all missing/empty entries from all entries in a column.\\
	ratio-alpha-chars & The average of the proportion of alphabetic characters from each entry.\\
	\rowcolor{Gray}
	prop-numerical-chars & The average of the proportion of numeric characters from each entry.\\
	prop-whitespace-chars    & The average of the proportion of whitespace characters from each entry.\\
	\rowcolor{Gray}
	prop-entries-with-at-sign& The proportion of entries with an '@' sign. \\
	prop-entries-with-hyphen & The proportion of entries with a '-'. \\
	\rowcolor{Gray}
	prop-range-format        & The proportion of entries which follow some numerical range format   ([0-9]+)-([0-9]+) \\
	is-discrete              & A binary indicator if the column entries are discrete. \\
	\rowcolor{Gray}
	entropy-for-discrete-values & The entropy of the entries of a column.\\
	shannon-entropy          & Shannon's entropy for the vocabulary specified in the description of the feature 'char-dist-features'\\
	\end{tabular}
\end{table}

@@

	\begin{tabular}{p{2.5cm}p{5cm}}
	Feature name & Description \\ \hline
inferred-data-type      & Produces a vector which contains the inferred data type.  
Only a single element is non-zero in the vector (1-of-K format).  
The set of inferred types are in float, integer, long, bool, date, time, datetime, string. \\
	\rowcolor{Gray}
char-dist-features      & Produces a vector for character distribution of column values;
considered characters are printable characters
%                        0123456789abcdefghijklmnopqrstuvwxyz
%                        ABCDEFGHIJKLMNOPQRSTUVWXYZ!"
%                        #$%&\'()*+,-./:;<=>?@[\\]^_`{|}~ \t\n\r\x0b\x0c
\\
stats-of-text-length    & Computes the mean, median, mode, min, max of string lengths from a columns entries.\\
	\rowcolor{Gray}
stats-of-numerical-type & Computes the mean, median, mode, min, max of entries from a column of numerical type.\\

prop-instances-per-class-in-knearestneighbours         & This finds the $k$ columns from the training set with most similar names.
It then produces a vector of class proportions from these $k$ neighbours.
Parameters:
"num-neighbours": The number of neighbours to find ($k$).\\
	\rowcolor{Gray}
mean-character-cosine-similarity-from-class-examples   & For some instance $t$, computes the character distribution 
from this instance's entries, and produces a vector of the average
cosine similarity of $t$ with training instances of each class.\\
min-editdistance-from-class-examples                   & For some instance $t$, produces a vector of the minimum column name
edit distance between $t$ and all training instances of each class.
Parameters:
``cache-size'': size of cache to store edit distances of words;
``max-comparisons-per-class'' : max number of training instances to compare for each class.\\

	\rowcolor{Gray}
min-wordnet-jcn-distance-from-class-examples            & For some instance $t$, produces a vector of the minimum column name
JCN WS4J distance between $t$ and all training instances of each class.
Parameters:
"cache-size": size of cache to store wordnet distances of words; 
"max-comparisons-per-class": max number of training instances to compare
against for each class.\\

min-wordnet-lin-distance-from-class-examples            & For some instance $t$, produces a vector of the minimum column name
LIN WS4J distance between $t$ and all training instances of each class.
Parameters:
"cache-size": size of cache to store wordnet distances of words; 
"max-comparisons-per-class": max number of training instances to compare
against for each class.\\

@@

\clearpage\newpage
\subsection{\minizinc{} Model \label{Ann:MZ}}
\renewcommand{\lstlistingname}{Model}
\begin{minipage}{\textwidth}
%\begin{figure}[!h]
	\begin{lstlisting}[caption={\minizinc{} model (part 1 of 2)},captionpos=b]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% MiniZinc model to solve the Rel2Onto problem             %%
%%  as a composition of Steiner Tree Problem                %%
%%  and a matching problem                                  %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%% Steiner Tree Problem
int: nbV;                               % Nb of nodes in alignment graph
int: nbE;                               % Nb of edges in alignment graph
set of int: nodes = 1..nbV;    
set of int: edges = 1..nbE;
set of int: cnodes;                     % Class nodes in alignment graph
set of int: dnodes;                     % Data nodes in alignment graph
array [nodes] of var bool: vs;          % Is a node of the alignment graph in the solution?
array [edges] of var bool: es;          % Is an edge of the alignment graph in the solution?
array [nodes,edges] of bool: adjacent;  % adjacent[n,e] <=> node n is an endnode of edge e
array [edges] of nodes: heads;          % Endpoint 1 of each edge (the graph is undirected)
array [edges] of nodes: tails;          % Endpoint 2 of each edge
array[1..2,edges] of int: endpairs = array2d(1..2,edges,tails++heads);  % Endpoints of each edge
array [edges] of int: ws;               % Weights of edges in alignment graph
var 0..sum(ws): w;                      % Weight of the solution corresponding to the alignment part
constraint steiner_tree(vs,es,adjacent,endpairs,w,ws);  % Steiner Tree constraint
constraint forall (d in dnodes) 
                   (vs[d] <-> sum(e in edges where adjacent[d,e])(es[e]) == 1);% Only 1 edges per data node

%%%%%%% Matching Problem
int: nbA;                                           % Nb of attribute nodes
set of int: atts = 1..nbA;                 
array [atts] of set of dnodes: attribute_domains;   % Data nodes to which each attribute can attach (usually all) 
array [atts] of var dnodes: match;                  % Match between Attributes and Data nodes
array[atts,dnodes] of int: match_costs;             % Weight of match edges
constraint forall (a in atts) (match[a] in (attribute_domains[a]));   % Must match someone in the domain
constraint forall (a in atts) (vs[match[a]]);       % The data node matched must be in the subgraph of the alignment
constraint alldifferent(match);                     % Exact match
var 0..sum(a in atts)(sum(n in dnodes)(match_costs[a,n])): wm;  % Total cost from matching part :
constraint wm = sum(a in atts) (match_costs[a,match[a]]);

	\end{lstlisting}
%	\caption{\minizinc{} model (1/2).}
%\end{figure}
\end{minipage}
\clearpage
\newpage
\begin{minipage}{\textwidth}
%\begin{figure*}[t]
	\begin{lstlisting}[caption={\minizinc{} model (part 2 of 2)},captionpos=b]
%%%%%%% Unknown nodes
int: nb_unknown_nodes;                 % Number of unknown data nodes
set of int: unknowns = 1..nb_unknown_nodes;
array [int] of int: all_unk_nodes;     % List of all unknown nodes (first two elements are class nodes, by convention)
int: node_root = all_unk_nodes[1];     % "Root" class node
int: node_unk = all_unk_nodes[2];      % "Unknown" class node
constraint if nb_unknown_nodes > 0 then
    vs[node_unk] <-> vs[node_root]     % If class node "Unknown" is used, class node "Root" is used as well
 /\ vs[node_root] <-> sum(e in edges where adjacent[node_root,e])(es[e]) in {1,2}
else true %Ignore constraint
endif;

%%%%%%% Patterns
int: nb_patterns;                     % Number of patterns
set of int: PATTS = 1..nb_patterns;
array[PATTS] of int: patterns_w;      % Support of patterns
array[PATTS] of set of int: patterns; % The actual patterns
array[PATTS] of var bool: c_p;        % Is pattern chosen to be part of the solution?
constraint forall (p in PATTS) ((sum (e in patterns[p])   % All edges of a pattern are used <-> pattern is used
							                  	(es[e]) == card(patterns[p])) <-> c_p[p]);
var 0..sum(patterns_w): wp;           % Total prize collected from patterns : 
constraint if nb_patterns = 0 then wp = 0 else wp = sum(p in PATTS) (c_p[p] * patterns_w[p]) endif;

%%%%%%% Final objective function 
var int: obj = w + wm - wp;           % Total = (cost of semantic model) - (prize collected from patterns)

@@

ann: cheap_edges_fill_pattern;        % Custom search strategy for patterns (implemented in Chuffed)
ann: custom(ann, array[int] of int, array[int] of var bool);
array[int] of ann: cefp = 
if nb_patterns == 0 then []           % This builds a list in a format understood by Chuffed, to work on the "es" vars 
else [custom(cheap_edges_fill_pattern,[nbE]++[nb_patterns]++ws++[ card(patterns[p]) | p in PATTS]++patterns_w++array1d([ e | p in PATTS, e in patterns[p]]),es)] endif;

% The solver will use these strategies in order:
% 1) Try using the cheapest matches, until a complete match is found (i.e. for all attributes)
% 2) If there are patterns, use a specific search called "cheap_edges_fill_pattern". (Implemented in Chuffed only)
%    It will start adding edges in increasing order of cost, while trying to fill patterns that are "worth" filling.
%    E.g. We already set es[x] = true for x in {e1,e2,e3}; if adding ws[e5] = 4 and ws[e7] = 6, but
%    {e1,e2,e3,e7} form a pattern p of patterns_w[p] = 3, then ws[e7] - patterns_w[p] <  ws[e5], so e7 is selected first
%    If there are no patterns, this strategy is ignored.
% 3) Try to eliminate edges by decreasing cost: it will eliminate the more expensive edges first, until
%    the strictly necesary cheap edges are added by the propagators. This is used only when there are no 
%    patterns, otherwise (2) already selects the edges.
